How to Analyze Build Logs---A~Comparative Study of Chunk~Retrieval~Techniques

### 2020-04-16 tse-paper/content.tex
write more in lit sur discussion, relates to #40

\begin{figure}[hb]
	\centering
	\includegraphics[width=\columnwidth, clip]{img/lit_survey.pdf}
	\caption{Our literature selection process, following
	\cite{petersen2015guidelines}.}
	\label{fig:lit-survey}
\end{figure}

@@

In the following section, we explain how build log analysis differs
from system log analysis.
Further, we describe how we selected the articles for our literature
survey and the data extraction from the resulting articles.
We present our results on how buildlogs are analyzed in literature
and discuss the implecations of these findings on our research.

\lstset{
  morekeywords={INFO, WARN, 2008, 11, 09},
  alsoletter=-20819,
  keywordstyle=\bfseries\color{Plum},
  escapeinside=**
}
\begin{figure*}[!tb]
  \centering
  \lstinputlisting[breaklines=true]{listings/syslog.txt}
  \caption{System Log excerpt.
Example adapted from~\cite{he2017towards}.}
  \label{lst:system-log}
\end{figure*}

@@

The techniques developed for system log analysis can also be applied
to build logs.
One example is comparing execution traces to reference
traces of intended behavior to detect anomalies.
Amar et
al.~\cite{amar2019mining} employed a similar approach to detect
relevant lines in build logs.
However, many of the techniques developed to analyze system logs
leverage their inherent event structure.
As build logs lack this sturcture, we focussed our literature
survey on articles that as a source of information.
% In this article, we focus on extracting a single specified information
% from the build log as a whole with chunk retrieval techniques.
% Chunk
% retrieval techniques are used as a part of log parsing to retrieve the
% values of variable parts in a log message, e.g.\ by using regular
% expressions~\cite{nagappan2010abstracting,xu2009detecting}.

\subsection{Study Selection}
\Cref{fig:lit-survey} presents an overview of our study selection
process.
Since the field of build log analysis is so young, we
surveyed a population of scientific material as broad as possible.
We did not limit our sources to specific Software Engineering venues, as
is often done~\cite{petersen2015guidelines}, but survey all sources
monitored by Google Scholar, including Bachelor's and Master's
theses and academic slide decks.
We use Google Scholar as opposed to Scopus or
other databases, as it is the search engine for acdemic articles with
the broadest and most current index, including preprints.

@@

First, we filtered articles based on (1) their title and (2) the
contextual information Google Scholar displayed.
We inlcuded only
publications written in a language intelligible to the authors (that
is, English or German), which disregarded fewer than 1\% of search
hits.
We then unified
the results of the four search queries based on the link as the
identifying element, removing 60 articles that appeared in more than
one search query.
We removed further duplicates based on the title and were
left with 256 (16\%) of the original search results.
% removed duplicates of data extraction pass already here

@@

We wanted to characterize how researchers use buildlogs as input to
any processes.
We were interesed in
\begin{itemize}
  \item what information they are retrieving from the logs and
  what they use the retrieved information for (\textbf{RQ1.1})
  \item which technique they are using to process the build logs,
  in how much detail it is described and whether they published their
  tools (\textbf{RQ1.2})
  \item what kind of build logs they are using and the origin of these
  logs.
(\textbf{RQ1.3})
\end{itemize}
To objecifiy our data extraction we created a template of the targeted
factors and performed a pilot, as reccommended by Petersen et
al.~\cite{petersen2015guidelines}.
Two authors extracted the data from five articles and we held a
consensus meeting to unify our understanding
of the data extraction template.
The remaining articles were divided over these two authors.
We inspected the articles as closely as necessary,
starting from occurences of ``log'' or ``build'' within the text.

@@

\begin{figure*}[tbhp]
		\centering
		\includegraphics[width=0.7\textwidth, trim={1.1cm 0.4cm
		1.5cm 0.5cm},
		clip]{img/lit-sur/log_producer_annotated.pdf}
		\caption{Frequency of supported log producers.}
		\label{fig:litsur:log_producer}
\end{figure*}

\subsubsection{Retrieved Information}
The information that was targeted varied from article to article.
Most prominent was the search for errors (39\%) which occured within
the build process, as presented in \Cref{fig:litsur:info_target}.
This is followed by extracting the executed or the failing
tests (both 11\%).
There were also numerous cases which targeted more specialized
information, such as the environments used during the
build~\cite{zolfagharinia2017not}, packages
installed~\cite{selberg2012use} or hints to source files which are
related to build failures~\cite{ren2018automated}.

In most cases (51\%), a chunk was retrieved from the log.
Build logs were scanned for compilation
errors~\cite{clemencic2014new} or
the duration of tasks within the build~\cite{zhang2016android}.
36\% of the articles classified the build logs into categories.
For instance distinguishing failures caused by developers from
failures casued
by infrastructure errors~\cite{lindqvist2019detection} or
whether projects use automated statc analysis
tools~\cite{kavaler2019tool}.
Other types of information (e.g.\ counting
lines, summarization) were only retrieved
in one case each.

The majority of articles (46\%) used the outcome of their build log
analysis for research purpopses, as shown in \Cref{fig:litsur:use}.
Further uses were giving feedback to developers (10\%) or automatic
on-ward processing of the retrieved information (8\%).

@@

The most mentioned method (34\%) was using a parser, where we also
included
rather unprecise statements like ``we parse the build logs''.
25\% of the articles described the use of regular expressions and 20\%
inspected the logs manually.
Another 8\% employed machine learning techniques such as natural language
processing or information retrieval techniques.
To give examples,
Seo et al.~\cite{seo2014programmers} developed a custom
parser to classify error messages, while Vassallo et
al.~\cite{vassallo2017a-tale} analyzed build logs with regular
expressions.
Ghaleb et al.~\cite{ghaleb2019studying} used a compound approach.
They start with manual categorization of the build logs and select
search strings that identify their targeted categories.
Based on the search strings they create a script which automatically
classifies the remaining logs.

@@

% Seo et al.~\cite{seo2014programmers}
% They based their analyses on sets of
% build logs collected from industry partners.
% They	reported by Java and C++ builds.

% Rausch et al.~\cite{rausch2017empirical} analyze CI builds of open
% source Java projects

% Vassallo et al.~\cite{vassallo2017a-tale} compare open source projects
% in Java to industrial ones.
% Their data stems from analyzing
% build logs from \emph{TravisTorrent}~\cite{beller2017travistorrent}
% with regular expressions.

% Vassallo et
% al.~\cite{vassallo2018un-break}
% They parse
% Maven~\cite{maven2019website} build logs into a structured
% representation and create hint generators.
% For example, one of the hint generators queries stack overflow for
% discussions related to why the build failed.

@@

\subsection{Discussion}
Our literature survey sheds light on how build log analysis is used within
computer science research.
In this section we will summarize our findings and relate them to the
approach we are taking to improve build log analysis with chunk
retrieval techniques.

@@

We have observed very little reuse of build log analysis tools.
This can stem from the fact that publishing replication packages and
research prototypes became popular only in recent years or from the
high specialization of the developed tools in regards to supported
build environments and targeted information.
Several articles noted the variety in log formats for different build
environments, which requires customizing build log techniques for
every supported build environment.
We implement build environment agonostic build log analysis
techniques, which are customized to a specific project by
the training examples which the user provides.
In addition, the \emph{LogChunks}~\cite{brandt2020logchunks}
data set we chose for our evaluation covers a broad range of build
environments.
Further, the training examples also specify the log chunk targeted by the
retrieval, which enable our chunk retrieval techniques to retrieve a
broad range of information from build logs.

Urli et al.~\cite{urli2018design} strongly discouraged from parsing
build logs for information as it is ``too error prone'', other
articles pointed to the effort of developing a custom tool.
Regular expressions are known to be tedious to
maintain~\cite{michael2019regexes}.
For this reason, we are focussing % TODO: less development time,
% only giving examples

This task of retrieving specific chunks of text from the
build logs can be solved by the chunk retrieval techniques we compare
in this article.
Our results can support researchers in choosing a
suitable technique for their data set of build logs and the chunks
they want to retrieve.
By relieving them from building custom parsers
we enable them to cover a much wider range of languages and build
tools in their studies.

% \subsubsection{Retrieved Information}
% why from build logs? not avaliable in cleaner format much use for
% research, a bit of developer information / tools following: research
% so many logs that unfeasible to do manually, less focus on tools for
% devs (caveat! we excluded books and documentation on industrial tools
% -> many approaches there to make build logs digestible, tough many
% focus on general minimizing of the data presented, not retrieving
% specific chunks)

% \subsubsection{Retrieval Technique}
% broadly mentioned shortly, rarely explained in detail we follow that
% researchers often see this as an unimportant side fact work needed to
% be done to get the big work done

% \subsubsection{Build Logs}
% often focussed on specific logs, sometimes mentioned because of effort
% developing a parser other times also because other tools used in
% further research are limiting the scope (reference paper with PIT
% tool).
% many focus on maven because of it's prevalence also in respect
% to other tools much research done on travis ci logs b/c of their
% abundance and open availability
\section{Chunk Retrieval Techniques}
\label{sec:techniques}
This section introduces key concepts of the three chunk retrieval
techniques we study in this article, namely program synthesis by
example (PBE), common text similarity (CTS), and keyword search (KWS).

@@

\subsection{Characteristics of Chunk Retrieval Techniques}
\label{sec:blirt}
In this article, we evaluate different techniques to automatically
retrieve pieces of information that appear literally in the build
logs.
%, i.e., techniques that do not aggreagte, combine, or deducted
%information.
We call such pieces of information from build logs
\emph{chunks}, and the techniques \emph{chunk retrieval techniques}.
The techniques we investigate here do not require a formal lexer and
parser to analyze the entire structure of build logs, but focus on
ad-hoc extracting just one specific piece of information per
configuration.

With the term \textit{configuration}, we abstract over the training
and parametrization that different techniques require in different
forms.
A configuration can be explicitly stated or implicitly dervied
by learning through provided examples.
It is therefore a manual
specification of which information the chunk retrieval should target.
It also supplies the necessary information for the technique to
identify the targeted information chunk in a build log.
Each chunk
retrieval technqiue has a specific \textit{granularity}, i.e.,\ the
smallest piece of text it can return (e.g., a line, or a word).
The
granularity might be adjustable by configuration.
\emph{Running a
chunk retrieval technique} means to execute a fully configured
technique to consumes as input a build log in plain text format and to
produce an array as output.
The array consists of substrings of the
build log text.

\begin{table*}[]
\centering
\caption{Chunk retrieval techniques.}
\begin{tabularx}{\textwidth}{@{}XlXlXX@{}}
\toprule
Name			     & Acronym & Identification Technique
& Granularity & Configuration & Source		  \\
\midrule
Program Synthesis by Example & PBE     & Regular expression program
& Character   & In/output examples	\\
Common Text Similarity	     & CTS     & TF-IDF \& cosine similarity,
expected number of lines & Line        & Output examples	   \\
Keyword Search		     & KWS     & Keywords, expected number of
lines			 & Line        & Keywords, context length  \\
Random Line Retrieval	     & RLR     & Random sample
& Line	      & Retrieval length	  \\
\bottomrule
\end{tabularx}
\label{tab:ctr}
\end{table*}

\subsection{Program Synthesis by Example (PBE)}
\label{sec:expl-pbe}
The concept of \emph{Programming by Example} aims to capture the
intent of the user through examples which they provide.
Leveraging this apprach, we implemented a chunk retrieval technique
for build logs.
Our implementation is based and highly influenced by the
\emph{Microsoft PROSE library}~\cite{prose2019webpage}.
The PROSE library is powered by the generic program synthesis framework
\emph{FlashMeta}~\cite{polozov2015flashmeta:} and the specialized
text extraction DSL \emph{FlashExtract}~\cite{le2014flashextract:}.
Both enable us to synthesize regular expression programs
consistent~\cite{mitchell1982generalization} with a set of in-
and output examples given by the user.
The usage of example enables the user to configure a chunk
retrieval without understanding the whole structure of
the build log.

@@

\subsection{Random Line Retrieval}
\label{sec:expl-rlr}
In our evaluation, we want to compare against a baseline of randomly
picking lines from the build log.
RLR mimicks the
situation of guessing blindly which lines might be interesting.
The only configuration option for RLR is the number of lines it should
return.
For a fair comparison to the other techniques (whose number of
returned lines is dynamic), we configure RLR to return the average
number of lines in the chunks of the training examples.

>>>>

\begin{figure}[htbp]
	\centering
	\includegraphics[width=\columnwidth, clip]{img/lit_survey.pdf}
	\caption{Our literature selection process, following
	\cite{petersen2015guidelines}.}
	\label{fig:lit-survey}
\end{figure}

@@

In the following section, we explain how build log analysis differs
from system log analysis.
Further, we describe how we selected the articles for our literature
survey and the data extraction from the resulting articles.
We present our results on how build logs are analyzed in literature
and discuss the implications of these findings on our research.

\lstset{
  morekeywords={INFO, WARN, 2008, 11, 09},
  alsoletter=-20819,
  keywordstyle=\bfseries\color{Plum},
  escapeinside=**
}
\begin{figure*}[htbp]
  \centering
  \lstinputlisting[breaklines=true]{listings/syslog.txt}
  \caption{System Log excerpt.
Example adapted from~\cite{he2017towards}.}
  \label{lst:system-log}
\end{figure*}

@@

The techniques developed for system log analysis can also be applied
to build logs.
One example is comparing execution traces to reference
traces of intended behavior to detect anomalies.
Amar et
al.~\cite{amar2019mining} employed a similar approach to detect
relevant lines in build logs.
However, many of the techniques developed to analyze system logs
leverage their inherent event structure.
As build logs lack this structure, we focussed our literature
survey on articles that as a source of information.
% In this article, we focus on extracting a single specified information
% from the build log as a whole with chunk retrieval techniques.
% Chunk
% retrieval techniques are used as a part of log parsing to retrieve the
% values of variable parts in a log message, e.g.\ by using regular
% expressions~\cite{nagappan2010abstracting,xu2009detecting}.

\subsection{Study Selection}
\Cref{fig:lit-survey} presents an overview of our study selection
process.
Since the field of build log analysis is so young, we
surveyed a population of scientific material as broad as possible.
We did not limit our sources to specific Software Engineering venues, as
is often done~\cite{petersen2015guidelines}, but survey all sources
monitored by Google Scholar, including Bachelor's and Master's
theses and academic slide decks.
We use Google Scholar as opposed to Scopus or
other databases, as it is the search engine for academic articles with
the broadest and most current index, including preprints.

@@

First, we filtered articles based on (1) their title and (2) the
contextual information Google Scholar displayed.
We included only
publications written in a language intelligible to the authors (that
is, English or German), which disregarded fewer than 1\% of search
hits.
We then unified
the results of the four search queries based on the link as the
identifying element, removing 60 articles that appeared in more than
one search query.
We removed further duplicates based on the title and were
left with 256 (16\%) of the original search results.
% removed duplicates of data extraction pass already here

@@

We wanted to characterize how researchers use build logs as input to
any processes.
We were interested in
\begin{itemize}
  \item what information they are retrieving from the logs and
  what they use the retrieved information for (\textbf{RQ1.1})
  \item which technique they are using to process the build logs,
  in how much detail it is described and whether they published their
  tools (\textbf{RQ1.2})
  \item what kind of build logs they are using and the origin of these
  logs.
(\textbf{RQ1.3})
\end{itemize}
To objectifiy our data extraction we created a template of the targeted
factors and performed a pilot, as recommended by Petersen et
al.~\cite{petersen2015guidelines}.
Two authors extracted the data from five articles and we held a
consensus meeting to unify our understanding
of the data extraction template.
The remaining articles were divided over these two authors.
We inspected the articles as closely as necessary,
starting from occurrences of ``log'' or ``build'' within the text.

@@

\begin{figure}[tbhp]
		\centering
		\includegraphics[width=\columnwidth, trim={1.1cm 0.4cm
		1.5cm 0.5cm},
		clip]{img/lit-sur/log_producer_annotated.pdf}
		\caption{Frequency of supported log producers.}
		\label{fig:litsur:log_producer}
\end{figure}

\subsubsection{Retrieved Information}
The information that was targeted varied from article to article.
Most prominent was the search for errors (39\%) which occurred within
the build process, as presented in \Cref{fig:litsur:info_target}.
This is followed by extracting the executed or the failing
tests (both 11\%).
There were also numerous cases which targeted more specialized
information, such as the environments used during the
build~\cite{zolfagharinia2017not}, packages
installed~\cite{selberg2012use} or hints to source files which are
related to build failures~\cite{ren2018automated}.

In most cases (51\%), a chunk was retrieved from the log.
Build logs were scanned for compilation
errors~\cite{clemencic2014new} or
the duration of tasks within the build~\cite{zhang2016android}.
36\% of the articles classified the build logs into categories.
For instance distinguishing failures caused by developers from
failures caused
by infrastructure errors~\cite{lindqvist2019detection} or
whether projects use automated static analysis
tools~\cite{kavaler2019tool}.
Other types of information (e.g.\ counting
lines, summarization) were only retrieved
in one case each.

The majority of articles (46\%) used the outcome of their build log
analysis for research purposes, as shown in \Cref{fig:litsur:use}.
Further uses were giving feedback to developers (10\%) or automatic
on-ward processing of the retrieved information (8\%).

@@

The most mentioned method (34\%) was using a parser, where we also
included
rather imprecise statements like ``we parse the build logs''.
25\% of the articles described the use of regular expressions and 20\%
inspected the logs manually.
Another 8\% employed machine learning techniques such as natural language
processing or information retrieval techniques.
To give examples,
Seo et al.~\cite{seo2014programmers} developed a custom
parser to classify error messages, while Vassallo et
al.~\cite{vassallo2017a-tale} analyzed build logs with regular
expressions.
Ghaleb et al.~\cite{ghaleb2019studying} used a compound approach.
They start with manual categorization of the build logs and select
search strings that identify their targeted categories.
Based on the search strings they create a script which automatically
classifies the remaining logs.

@@

@@

\subsection{Discussion}
Our literature survey sheds light on how build log analysis is used within
computer science research.
Hereafter we will summarize our findings and relate them to the
approach we are taking to improve build log analysis with chunk
retrieval techniques.

@@

We have observed very little reuse of build log analysis tools.
This can stem from the fact that publishing replication packages and
research prototypes became popular only in recent years or from the
high specialization of the developed tools in regards to supported
build environments and targeted information.
Several articles noted the variety in log formats for different build
environments, which requires customizing build log techniques for
every supported build environment.
\Cref{fig:litsur:log_producer} shows that few build environments are
targeted by many studies and tools (A).
The majority of build environments was however only targeted by one
article each (B).
From this we follow that, while custom parsers are beneficial for the
small group of environments targeted by many (A), all and especially
the rarer investigated environments (B) would benefit from a generic
solution.
We implement build environment agnostic build log analysis
techniques, which are customized to a specific project by
the training examples which the user provides.
In addition, the \emph{LogChunks}~\cite{brandt2020logchunks}
data set we chose for our evaluation covers a broad range of build
environments.
% Further, the training examples also specify the log chunk targeted
by the
% retrieval, which enable our chunk retrieval techniques to retrieve a
% broad range of information from build logs.

The articles within our study mainly tried to retrieve or classify
errors or reasons for the build to fail from the logs.
Therefore we chose the log chunk ``describing why the build failed''
~\cite{brandt2020logchunks} for our empirical comparison study
of chunk retrieval techniques.
The techniques we implement are in principle agnostic to the targeted
log chunk, thus we believe our results can be generalized to other
kinds of log chunks.

Urli et al.~\cite{urli2018design} strongly discouraged from parsing
build logs for information as it is ``too error prone'', other
articles pointed to the effort of developing a custom tool.
Regular expressions are known to be tedious to
maintain~\cite{michael2019regexes}.
For this reason, we are focussing on techniques that are configured
through examples consisting of logs and corresponding chunks.
These examples are provided by the user and can simply be
exchanged whenever new cases are introduced.

% This task of retrieving specific chunks of text from the
% build logs can be solved by the chunk retrieval techniques we compare
% in this article.
% Our results can support researchers in choosing a
% suitable technique for their data set of build logs and the chunks
% they want to retrieve.
% By relieving them from building custom parsers
% we enable them to cover a much wider range of languages and build
% tools in their studies.

In summary, this literature study shows the need for build log
analysis techniques that are build-environment agnostic, automated
and can be configured with small effort.
We achieve this by developing three techniques for which the
information to be retrieved is specified through examples or keywords.

We propose the notion of chunk retrieval techniques which retrieve
a substring, that describes a specific information, from a build log.
Generally these can also be used to classify build logs and therefore
cover the great majority of the type of log analysis we saw within
our survey.

We implement three chunk retrieval techniques based on approaches
described by several articles:
PBE mimics regular expressions and parsers while simplifying
their development.
We chose text similarity (CTS) as a common information retrieval
technique, and implement the ad-hoc approach of
searching for specific strings or keywords within KWS\@.

Our evaluation of the three chunk retrieval techniques is done on the
\emph{LogChunks} data set, because it encompasses a broad range of
build environments which enables us to measure the generic
applicability of our techniques.
The targeted chunks are describing error messages or
``the reason the build failed'', which was also the information
most targeted by the articles in our survey.

\section{Chunk Retrieval Techniques}
\label{sec:techniques}
This section introduces key concepts of the three chunk retrieval
techniques we study in this article, namely program synthesis by
example (PBE), common text similarity (CTS), and keyword search (KWS).

@@

\begin{table*}[htbp]
\centering
\caption{Chunk retrieval techniques.}
\begin{tabularx}{\textwidth}{@{}llXll@{}}
\toprule
Name			     & Acronym & Identification Technique
& Granularity & Configuration \\
\midrule
Program Synthesis by Example & PBE     & Regular expression program
& Character   & In/output examples	\\
Common Text Similarity	     & CTS     & TF-IDF \& cosine similarity,
expected number of lines & Line        & Output examples	   \\
Keyword Search		     & KWS     & Keywords, expected number of
lines			 & Line        & Keywords, context length  \\
Random Line Retrieval	     & RLR     & Random sample
& Line	      & Retrieval length	  \\
\bottomrule
\end{tabularx}
\label{tab:ctr}
\end{table*}

\subsection{Characteristics of Chunk Retrieval Techniques}
\label{sec:blirt}
In this article, we evaluate different techniques to automatically
retrieve pieces of information that appear literally in the build
logs.
%, i.e., techniques that do not aggregate, combine, or deducted
%information.
We call such pieces of information from build logs
\emph{chunks}, and the techniques \emph{chunk retrieval techniques}.
The techniques we investigate here do not require a formal lexer and
parser to analyze the entire structure of build logs, but focus on
ad-hoc extracting just one specific piece of information per
configuration.

With the term \textit{configuration}, we abstract over the training
and parametrization that different techniques require in different
forms.
A configuration can be explicitly stated or implicitly derived
by learning through provided examples.
It is therefore a manual
specification of which information the chunk retrieval should target.
It also supplies the necessary information for the technique to
identify the targeted information chunk in a build log.
Each chunk
retrieval technique has a specific \textit{granularity}, i.e.,\ the
smallest piece of text it can return (e.g., a line, or a word).
The
granularity might be adjustable by configuration.
\emph{Running a
chunk retrieval technique} means to execute a fully configured
technique to consumes as input a build log in plain text format and to
produce an array as output.
The array consists of substrings of the
build log text.

\subsection{Program Synthesis by Example (PBE)}
\label{sec:expl-pbe}
The concept of \emph{Programming by Example} aims to capture the
intent of the user through examples which they provide.
Leveraging this approach, we implemented a chunk retrieval technique
for build logs.
Our implementation is based and highly influenced by the
\emph{Microsoft PROSE library}~\cite{prose2019webpage}.
The PROSE library is powered by the generic program synthesis framework
\emph{FlashMeta}~\cite{polozov2015flashmeta:} and the specialized
text extraction DSL \emph{FlashExtract}~\cite{le2014flashextract:}.
Both enable us to synthesize regular expression programs
consistent~\cite{mitchell1982generalization} with a set of in-
and output examples given by the user.
The usage of example enables the user to configure a chunk
retrieval without understanding the whole structure of
the build log.

@@

\subsection{Random Line Retrieval}
\label{sec:expl-rlr}
In our evaluation, we want to compare against a baseline of randomly
picking lines from the build log.
RLR mimics the
situation of guessing blindly which lines might be interesting.
The only configuration option for RLR is the number of lines it should
return.
For a fair comparison to the other techniques (whose number of
returned lines is dynamic), we configure RLR to return the average
number of lines in the chunks of the training examples.
